---
title: "Mental Health Analysis: 2023-2026"
author: "Carson Trego"
date: "2026-01-12"
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    mainfont: "Calibri"
    monofont: "Consolas"
    fontsize: 0.2em
    code-overflow: wrap
editor: visual
execute:
  echo: false
  warning: false
---

## DESCRIPTION

Across the last three years, three surveys were taken using almost the exact same methodology. Dental school students were asked to take a set of mental health surveys, and 3 metrics were scored (PHQ9,BURNOUT MINI Z,GAD7) Ids were not taken, and given the size of the school, it is very likely that many of the participants are the same people, so a person may have taken the quiz as a first year in the first survey, a second year in the second survey, and a third year in the third survey, and we would not be able to tell.

## PREPARATION

We have separate files with separate schemes for how the data is recorded, and we need to modify the data to all be in the same format. This whole section is that process, but it contains no analysis or plotting. For information to put on the abstract, skip to the PLOTTING and ANALYSIS sections.

```{r imports and startup }
library(dplyr)
library(ggplot2)
library(readxl)
library(readr)

options(width = 60)


reSeed <- 1 # seed for reproducabilty
list.files()
```

```{r load data}
df_23_24_orig <- read_excel('doNotModifyJustRead\\2023-2024 CDCSV.xlsx')
df_24_25_orig <- read_excel('doNotModifyJustRead\\2024-2025 CDCSV.xlsx')
df_25_26_orig <- read_excel('doNotModifyJustRead\\2025-2026 CDCSV.xlsx')

head(df_23_24_orig)
print('-------------------------')
head(df_24_25_orig)
print('-------------------------')
head(df_25_26_orig)
```

```{r get name info}
print('===========')
colnames(df_23_24_orig)
colnames(df_24_25_orig)
colnames(df_25_26_orig)
```

```{r get col names}


# 23-24
names_ALL_23_24 <- colnames(df_23_24_orig[5:20]) # ALL
names_PHQ_23_24 <- colnames(df_23_24_orig[5:12]) # PHQ 8
names_BUR_23_24 <- colnames(df_23_24_orig[13:13]) # BURN
names_GAD_23_24 <- colnames(df_23_24_orig[14:20]) # GAD 7


# 24-25
names_ALL_24_25 <- colnames(df_24_25_orig[4:19]) # ALL
names_PHQ_24_25 <- colnames(df_24_25_orig[4:11]) # PHQ 8
names_BUR_24_25 <- colnames(df_24_25_orig[12:12]) # BURN
names_GAD_24_25 <- colnames(df_24_25_orig[13:19]) # GAD 7


# 25-26
names_ALL_25_26 <- colnames(df_25_26_orig[4:19]) # ALL
names_PHQ_25_26 <- colnames(df_25_26_orig[4:11]) # PHQ 8
names_BUR_25_26 <- colnames(df_25_26_orig[12:12]) # BURN
names_GAD_25_26 <- colnames(df_25_26_orig[13:19]) # GAD 7
```

```{python write ans, eval=FALSE}
"""
23-25
BURN
0 0
1 1
2 2
3 3
4 4 FOR SOME REASON, NO ONE ANSWERED 4


24-25
BURN
THIS IS THE ORDER GIVEN BY THE MINI Z Q3
0 I enjoy my school, I have no symptoms of burnout.
1 I am under stress, and don't always have as much energy as I did, but I don't feel burned out.
2 I am definitely burning out and have one or more symptoms of burnout e.g. emotional exhaustion.
3 The symptoms of burnout that I'm experiencing won't go away. I think about school frustrations a lot.
4 I feel completely burned out. I am the point where I may need to seek help.

ALL ELSE
THIS IS THE ORDER GIVEN BY THE PHQ 9 AND GAD
0 Not at all
1 Several days
2 More than half the days
3 Nearly every day


25-26
0 I enjoy my school, I have no symptoms of burnout.
1 I am under stress, and don't always have as much energy as I did, but I don't feel burned out.
2 I am definitely burning out and have one or more symptoms of burnout e.g. emotional exhaustion.
3 The symptoms of burnout that I'm experiencing won't go away. I think about school frustrations a lot.
4 I feel completely burned out. I am the point where I may need to seek help.

ALL ELSE
0 Not at all
1 Several Days 
1 Several days 
2 More than half the days
2 More then half the days
3 Nealry every day 
3 Nearly every day 

"""
```

```{r get possible answers}

if (FALSE){
  table(df_23_24_orig[12])#[5:20])
  
  table(unlist( df_23_24_orig[, 5:20]), useNA = "always" )
  table(unlist( df_24_25_orig[, 4:19]), useNA = "always" )
  table(unlist( df_25_26_orig[, 4:19]), useNA = "always" )
  
  
  table(unlist( df_23_24_orig[, 13:13]), useNA = "always" )
  table(unlist( df_24_25_orig[, 12:12]), useNA = "always" )
  table(unlist( df_25_26_orig[, 12:12]), useNA = "always" )
}





map_scheme <- c(
  "I enjoy my school, I have no symptoms of burnout." = 0,
  "I am under stress, and don't always have as much energy as I did, but I don't feel burned out." = 1,
  "I am definitely burning out and have one or more symptoms of burnout e.g. emotional exhaustion." = 2,
  "The symptoms of burnout that I'm experiencing won't go away. I think about school frustrations a lot." = 3,
  "I feel completely burned out. I am the point where I may need to seek help." = 4,
  "Not at all" = 0,
  "Several days" = 1,
  "More than half the days" = 2,
  "Nearly every day" = 3,
  "Several Days" = 1,
  "More then half the days" = 2,
  "Nealry every day" = 3,
  "0" = 0,
  "1" = 1,
  "2" = 2,
  "3" = 3,
  "4" = 4
)


#df_23_24_mapped <- df_23_24_orig %>%
#  mutate(across(5:20, ~ map_scheme[.x] ))
#table(unlist( df_23_24_mapped[, 5:20]), useNA = "always" )


df_23_24_mapped <- df_23_24_orig %>%
  mutate(across(5:20, ~ recode(as.character(.x), !!!map_scheme, .default = NA_real_ )))
table(unlist( df_23_24_mapped[, 5:20]), useNA = "always")



df_24_25_mapped <- df_24_25_orig %>%
  mutate(across(4:20, ~ recode(.x, !!!map_scheme, .default = NA_real_ )))
table(unlist( df_24_25_mapped[, 4:19]), useNA = "always")


df_25_26_mapped <- df_25_26_orig %>%
  mutate(across(4:19, ~ recode(.x, !!!map_scheme, .default = NA_real_)))
table(unlist( df_25_26_mapped[, 4:19]), useNA = "always" )


colnames(df_23_24_mapped[5:20])
colnames(df_23_24_mapped[names_PHQ_23_24])

df_23_24_mapped <- df_23_24_mapped %>%
  mutate( PHQ_SCORE = rowSums(across(all_of(names_PHQ_23_24) ), na.rm=FALSE)) %>%
  mutate( BUR_SCORE = rowSums(across(all_of(names_BUR_23_24) ), na.rm=FALSE)) %>%
  mutate( GAD_SCORE = rowSums(across(all_of(names_GAD_23_24) ), na.rm=FALSE))
  

df_24_25_mapped <- df_24_25_mapped %>%
  mutate( PHQ_SCORE = rowSums(across(all_of(names_PHQ_24_25) ), na.rm=FALSE)) %>%
  mutate( BUR_SCORE = rowSums(across(all_of(names_BUR_24_25) ), na.rm=FALSE)) %>%
  mutate( GAD_SCORE = rowSums(across(all_of(names_GAD_24_25) ), na.rm=FALSE))

df_25_26_mapped <- df_25_26_mapped %>%
  mutate( PHQ_SCORE = rowSums(across(all_of(names_PHQ_25_26) ), na.rm=FALSE)) %>%
  mutate( BUR_SCORE = rowSums(across(all_of(names_BUR_25_26) ), na.rm=FALSE)) %>%
  mutate( GAD_SCORE = rowSums(across(all_of(names_GAD_25_26) ), na.rm=FALSE))




df_23_24_mapped <- df_23_24_mapped %>%
  rename_with(~ paste0("PHQQ", seq_along(names_PHQ_23_24)), all_of(names_PHQ_23_24)) %>%
  rename_with(~ paste0("BURNQ", seq_along(names_BUR_23_24)), all_of(names_BUR_23_24)) %>%
  rename_with(~ paste0("GADQ", seq_along(names_GAD_23_24)), all_of(names_GAD_23_24))

df_24_25_mapped <- df_24_25_mapped %>%
  rename_with(~ paste0("PHQQ", seq_along(names_PHQ_24_25)), all_of(names_PHQ_24_25)) %>%
  rename_with(~ paste0("BURNQ", seq_along(names_BUR_24_25)), all_of(names_BUR_24_25)) %>%
  rename_with(~ paste0("GADQ", seq_along(names_GAD_24_25)), all_of(names_GAD_24_25))

df_25_26_mapped <- df_25_26_mapped %>%
  rename_with(~ paste0("PHQQ", seq_along(names_PHQ_25_26)), all_of(names_PHQ_25_26)) %>%
  rename_with(~ paste0("BURNQ", seq_along(names_BUR_25_26)), all_of(names_BUR_25_26)) %>%
  rename_with(~ paste0("GADQ", seq_along(names_GAD_25_26)), all_of(names_GAD_25_26))


df_23_24_mapped <- df_23_24_mapped %>%
  rename(D_YEAR = `What year in dental school are you?`)

df_24_25_mapped <- df_24_25_mapped %>%
  rename(D_YEAR = `What year in dental school are you?`)

df_25_26_mapped <- df_25_26_mapped %>%
  rename(D_YEAR = `What year in dental school are you?
Mark only one oval.`)


# --- COHORT  ---
calculate_cohort <- function(d_year_str, survey_year_str) {
  # GET BASE YEAR
  base_year <- as.numeric(substr(survey_year_str, 1, 4))
  
  # CONVERT YEAR TO NUM
  year_num <- case_when(
    grepl("1|first|one", tolower(d_year_str)) ~ 1,
    grepl("2|second|two", tolower(d_year_str)) ~ 2,
    grepl("3|third|three", tolower(d_year_str)) ~ 3,
    grepl("4|fourth|four", tolower(d_year_str)) ~ 4,
    TRUE ~ NA_real_
  )
  
  # CALCULATE GRADUATION YEAR
  cohort_year <- base_year + 5 - year_num
  
  return(paste("Class of", cohort_year))
}
# ---------------------------

cols_to_keep = cols_to_keep <- c("PHQ_SCORE", 
                                 "BUR_SCORE", 
                                 "GAD_SCORE",
                                 "PHQQ1",
                                 "PHQQ2",
                                 "PHQQ3",
                                 "PHQQ4",
                                 "PHQQ5",
                                 "PHQQ6",
                                 "PHQQ7",
                                 "PHQQ8",
                                 "BURNQ1",
                                 "GADQ1",
                                 "GADQ2",
                                 "GADQ3",
                                 "GADQ4",
                                 "GADQ5",
                                 "GADQ6",
                                 "GADQ7",
                                 "D_YEAR",
                                 "Timestamp"
                                 )
df_list <- list(
  "2023-2024" = df_23_24_mapped %>% 
    mutate(Timestamp = as.character(Timestamp)) %>%
    select(any_of(cols_to_keep)),
  "2024-2025" = df_24_25_mapped %>%
    mutate(Timestamp = as.character(Timestamp)) %>%
    select(any_of(cols_to_keep)),
  "2025-2026" = df_25_26_mapped %>%
    mutate(Timestamp = as.character(Timestamp)) %>%
    select(any_of(cols_to_keep))
)

df_allYears <- bind_rows(df_list, .id = "Year") %>%
  mutate(Cohort = mapply(calculate_cohort, D_YEAR, Year))

```

## PLOTTING (IMPORTANT)

Now that we have done all of the data preparation, we have a single file (df_allYears) with all of the information. We can use this for plotting and

```{r plotting}
# Filter out unknown cohorts
plot_data <- df_allYears %>% filter(!grepl("NA", Cohort))

phq_cohort <- ggplot(plot_data, aes(x = Year, y = PHQ_SCORE, fill = Cohort)) +
  geom_violin(trim = FALSE, alpha = 0.5) + 
  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) + 
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "black") +
  stat_summary(fun = mean, geom = "line", aes(group = 1), color = "black", linewidth = 1) +
  facet_wrap(~Cohort) +  # <--- CRITICAL CHANGE: Facet by Cohort
  theme_minimal() +
  labs(x = "Academic Year", 
       y = "PHQ-9 Score",
       title = "PHQ-9 Scores by Cohort",
       subtitle = "Higher scores indicate greater depression severity. Black Line = Mean Score per Class Each Year") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), legend.position = "none")

bur_cohort <- ggplot(plot_data, aes(x = Year, y = BUR_SCORE, fill = Cohort)) +
  geom_violin(trim = FALSE, alpha = 0.5) + 
  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) + 
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "black") +
  stat_summary(fun = mean, geom = "line", aes(group = 1), color = "black", linewidth = 1) +
  facet_wrap(~Cohort) +
  theme_minimal() +
  labs(x = "Academic Year", 
       y = "Burnout Score",
       title = "Burnout Scores by Cohort",
       subtitle = "Higher scores indicate greater burnout severity. Black Line = Mean Score per Class Each Year") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), legend.position = "none")

gad_cohort <- ggplot(plot_data, aes(x = Year, y = GAD_SCORE, fill = Cohort)) +
  geom_violin(trim = FALSE, alpha = 0.5) + 
  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) + 
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "black") +
  stat_summary(fun = mean, geom = "line", aes(group = 1), color = "black", linewidth = 1) +
  facet_wrap(~Cohort) +
  theme_minimal() +
  labs(x = "Academic Year", 
       y = "GAD-7 Score",
       title = "GAD-7 Scores by Cohort",
       subtitle = "Higher scores indicate greater anxiety severity. Black Line = Mean Score per Class Each Year") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), legend.position = "none")


ggsave("PHQ_Cohort_Square.png", plot = phq_cohort, width = 8, height = 8, units = "in", dpi = 300, bg = "white")
ggsave("BURN_Cohort_Square.png", plot = bur_cohort, width = 8, height = 8, units = "in", dpi = 300, bg = "white")
ggsave("GAD_Cohort_Square.png", plot = gad_cohort, width = 8, height = 8, units = "in", dpi = 300, bg = "white")
```

```{r noweight}



df_means <- df_allYears %>%
  group_by(Year) %>%
  summarise(
    PHQ_mean = mean(PHQ_SCORE, na.rm = TRUE),
    BUR_mean = mean(BUR_SCORE, na.rm = TRUE),
    GAD_mean = mean(GAD_SCORE, na.rm = TRUE)
  )



p1 <- ggplot(df_allYears, aes(x = Year, y = PHQ_SCORE, fill = Year)) +
  geom_violin(alpha = 0.5, trim = FALSE, bw=0.5) +
  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) + 
  geom_point(data = df_means, aes(y = PHQ_mean), shape = 18, size = 4, color = "red") +
  geom_line(data = df_means, aes(y = PHQ_mean, group = 1), color = "red", linewidth = 1) +
  theme_minimal() +
  labs(x = "Academic Year", y = "PHQ Score",
       title = "PHQ", subtitle = "Black Line = Mean")


p2 <- ggplot(df_allYears, aes(x = Year, y = BUR_SCORE, fill = Year)) +
  geom_violin(alpha = 0.5, trim = FALSE, bw=0.5) +
  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) + 
  geom_point(data = df_means, aes(y = BUR_mean), shape = 18, size = 4, color = "red") +
  geom_line(data = df_means, aes(y = BUR_mean, group = 1), color = "red", linewidth = 1) +
  theme_minimal() +
  labs(x = "Academic Year", y = "BURN Score",
       title = "BURN", subtitle = "Black Line = Mean")

p3 <- ggplot(df_allYears, aes(x = Year, y = GAD_SCORE, fill = Year)) +
  geom_violin(alpha = 0.5, trim = FALSE, bw=0.5) +
  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) + 
  geom_point(data = df_means, aes(y = GAD_mean), shape = 18, size = 4, color = "red") +
  geom_line(data = df_means, aes(y = GAD_mean, group = 1), color = "red", linewidth = 1) +
  theme_minimal() +
  labs(x = "Academic Year", y = "GAD Score",
       title = "GAD", subtitle = "Black Line = Mean")

# Display the plots
print(p1)
print(p2)
print(p3)
```

## RAW ANALYSIS

```{r model check}
# Filter for analysis
df_model <- df_allYears %>%
  filter(!grepl("NA", Cohort)) %>%
  mutate(
    Year_Num = case_when(
      Year == "2023-2024" ~ 0,
      Year == "2024-2025" ~ 1,
      Year == "2025-2026" ~ 2
    ),
    D_YEAR = as.factor(D_YEAR) 
  )


# does the Year effect differ by Cohort?
#phq_model_interaction <- lm(PHQ_SCORE ~ Year_Num * Cohort, data = df_model)
#summary(phq_model_interaction)
#anova(phq_model_interaction)

# Separate models per cohort 
cohorts <- unique(df_model$Cohort)
for (c in c("Class of 2026","Class of 2027") ) {
  cat("\n==================================\n")
  cat("ANALYSIS FOR:", c, "\n")
  sub_data <- df_model %>% filter(Cohort == c)
  
  # Only run if we have multiple years of data for this cohort
  if (length(unique(sub_data$Year)) > 1) {
    print("--PHQ--")
    mod <- lm(PHQ_SCORE ~ Year_Num+I(Year_Num^2), data = sub_data)
    print(summary(mod))
    plot(mod)
    print("--BUR--")
    mod <- lm(BUR_SCORE ~ Year_Num+I(Year_Num^2), data = sub_data)
    print(summary(mod))
    plot(mod)
    print("--GAD--")
    mod <- lm(GAD_SCORE ~ Year_Num+I(Year_Num^2), data = sub_data)
    print(summary(mod))
    plot(mod)
  } else {
    cat("Not enough longitudinal data for this cohort.\n")
  }
}
```

```{r within most recent year}
df_recent <- df_allYears %>% 
  filter(Year == "2025-2026") %>%
  mutate(D_YEAR = as.factor(D_YEAR))


k_phq <- kruskal.test(PHQ_SCORE ~ D_YEAR, data =df_recent)
lm_phq <- lm(PHQ_SCORE ~ D_YEAR, data =df_recent)
summary(lm_phq)
# plot(lm(PHQ_SCORE ~ D_YEAR, data =df_recent))

k_bur <- kruskal.test(BUR_SCORE ~ D_YEAR, data =df_recent)
lm_bur <- lm(BUR_SCORE ~ D_YEAR, data =df_recent)
summary(lm_bur)
# plot(lm(BUR_SCORE ~ D_YEAR, data =df_recent))

k_gad <- kruskal.test(GAD_SCORE ~ D_YEAR, data =df_recent)
lm_gad <- lm(GAD_SCORE ~ D_YEAR, data =df_recent)
summary(lm_gad)
# plot(lm(GAD_SCORE ~ D_YEAR, data =df_recent))

get_lm_p <- function(model) {
  if (is.null(summary(model)$fstatistic)) return(NA)
  f <- summary(model)$fstatistic
  pf(f[1], f[2], f[3], lower.tail = FALSE)
}

# Kruskal-Wallis
raw_p_kw <- c(
  PHQ = k_phq$p.value,
  BUR = k_bur$p.value,
  GAD = k_gad$p.value
)
adj_p_kw <- p.adjust(raw_p_kw, method = "holm")

# Linear Model (ANOVA)
raw_p_lm <- c(
  PHQ = get_lm_p(lm_phq),
  BUR = get_lm_p(lm_bur),
  GAD = get_lm_p(lm_gad)
)
adj_p_lm <- p.adjust(raw_p_lm, method = "holm")


print("--- KRUSKAL-WALLIS RESULTS (Holm Adjusted) ---")
print(data.frame(
  Outcome = names(raw_p_kw),
  Raw_P = round(raw_p_kw, 5),
  Adj_P = round(adj_p_kw, 5),
  Signif = ifelse(adj_p_kw < 0.05, "*", "")
))

print("--- LINEAR MODEL RESULTS (Holm Adjusted) ---")
print(data.frame(
  Outcome = names(raw_p_lm),
  Raw_P = round(raw_p_lm, 5),
  Adj_P = round(adj_p_lm, 5),
  Signif = ifelse(adj_p_lm < 0.05, "*", "")
))
```

```{r pairwise recent}

cat("\n=================================================\n")
cat("PAIRWISE COMPARISONS (2025-2026 Data)\n")
cat("Test: Pairwise Wilcoxon Rank Sum (Holm-Adjusted)\n")
cat("=================================================\n")

cat("\n--- PHQ-9 Pairwise ---\n")
print(pairwise.wilcox.test(df_recent$PHQ_SCORE, df_recent$D_YEAR, p.adjust.method = "holm"))

cat("\n--- Burnout Pairwise ---\n")
print(pairwise.wilcox.test(df_recent$BUR_SCORE, df_recent$D_YEAR, p.adjust.method = "holm"))

cat("\n--- GAD-7 Pairwise ---\n")
print(pairwise.wilcox.test(df_recent$GAD_SCORE, df_recent$D_YEAR, p.adjust.method = "holm"))
```



```{r all year stats}

df_all_years_analysis <- df_allYears %>% 
  mutate(D_YEAR = as.factor(D_YEAR))

k_phq_all <- kruskal.test(PHQ_SCORE ~ D_YEAR, data = df_all_years_analysis)
lm_phq_all <- lm(PHQ_SCORE ~ D_YEAR, data = df_all_years_analysis)
summary(lm_phq_all)

k_bur_all <- kruskal.test(BUR_SCORE ~ D_YEAR, data = df_all_years_analysis)
lm_bur_all <- lm(BUR_SCORE ~ D_YEAR, data = df_all_years_analysis)
summary(lm_bur_all)

k_gad_all <- kruskal.test(GAD_SCORE ~ D_YEAR, data = df_all_years_analysis)
lm_gad_all <- lm(GAD_SCORE ~ D_YEAR, data = df_all_years_analysis)
summary(lm_gad_all)


# Kruskal-Wallis (All Years)
raw_p_kw_all <- c(
  PHQ = k_phq_all$p.value,
  BUR = k_bur_all$p.value,
  GAD = k_gad_all$p.value
)
adj_p_kw_all <- p.adjust(raw_p_kw_all, method = "holm")

# Linear Model (ANOVA) (All Years)
raw_p_lm_all <- c(
  PHQ = get_lm_p(lm_phq_all),
  BUR = get_lm_p(lm_bur_all),
  GAD = get_lm_p(lm_gad_all)
)
adj_p_lm_all <- p.adjust(raw_p_lm_all, method = "holm")


print("--- KRUSKAL-WALLIS RESULTS (All Years Combined, Holm Adjusted) ---")
print(data.frame(
  Outcome = names(raw_p_kw_all),
  Raw_P = round(raw_p_kw_all, 5),
  Adj_P = round(adj_p_kw_all, 5),
  Signif = ifelse(adj_p_kw_all < 0.05, "*", "")
))

print("--- LINEAR MODEL RESULTS (All Years Combined, Holm Adjusted) ---")
print(data.frame(
  Outcome = names(raw_p_lm_all),
  Raw_P = round(raw_p_lm_all, 5),
  Adj_P = round(adj_p_lm_all, 5),
  Signif = ifelse(adj_p_lm_all < 0.05, "*", "")
))
```

```{r pairwise all years}
cat("\n=================================================\n")
cat("PAIRWISE COMPARISONS (All Years Combined)\n")
cat("Test: Pairwise Wilcoxon Rank Sum (Holm-Adjusted)\n")
cat("=================================================\n")

cat("\n--- PHQ-9 Pairwise ---\n")
print(pairwise.wilcox.test(df_all_years_analysis$PHQ_SCORE, df_all_years_analysis$D_YEAR, p.adjust.method = "holm"))

cat("\n--- Burnout Pairwise ---\n")
print(pairwise.wilcox.test(df_all_years_analysis$BUR_SCORE, df_all_years_analysis$D_YEAR, p.adjust.method = "holm"))

cat("\n--- GAD-7 Pairwise ---\n")
print(pairwise.wilcox.test(df_all_years_analysis$GAD_SCORE, df_all_years_analysis$D_YEAR, p.adjust.method = "holm"))
```

```{r all year plotting}



# Combine all years but group by Grade Level (D_YEAR)
df_grade_means <- df_allYears %>%
  group_by(D_YEAR) %>%
  summarise(
    PHQ_mean = mean(PHQ_SCORE, na.rm = TRUE),
    BUR_mean = mean(BUR_SCORE, na.rm = TRUE),
    GAD_mean = mean(GAD_SCORE, na.rm = TRUE)
  )


p_phq_all <- ggplot(df_allYears, aes(x = D_YEAR, y = PHQ_SCORE, fill = D_YEAR)) +
  geom_violin(alpha = 0.5, trim = FALSE, bw = 0.5) +
  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) + 
  geom_point(data = df_grade_means, aes(y = PHQ_mean), shape = 18, size = 4, color = "black") +
  geom_line(data = df_grade_means, aes(y = PHQ_mean, group = 1), color = "black", linewidth = 1) +
  theme_minimal() +
  labs(x = "Grade Level", y = "PHQ-9 Score",
       title = "PHQ-9 Scores by Grade Level (All Years Combined)", subtitle = "Higher scores indicate greater depression severity. Black Line = Mean Score per Grade") +
  theme(legend.position = "none",
        plot.subtitle = element_text(size = 10))


p_burn_all <- ggplot(df_allYears, aes(x = D_YEAR, y = BUR_SCORE, fill = D_YEAR)) +
  geom_violin(alpha = 0.5, trim = FALSE, bw = 0.5) +
  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) + 
  geom_point(data = df_grade_means, aes(y = BUR_mean), shape = 18, size = 4, color = "black") +
  geom_line(data = df_grade_means, aes(y = BUR_mean, group = 1), color = "black", linewidth = 1) +
  theme_minimal() +
  labs(x = "Grade Level", y = "Burnout Score",
       title = "Burnout Scores by Grade Level (All Years Combined)", subtitle = "Higher scores indicate greater burnout severity. Black Line = Mean Score per Grade") +
  theme(legend.position = "none",
        plot.subtitle = element_text(size = 10))


p_gad_all <- ggplot(df_allYears, aes(x = D_YEAR, y = GAD_SCORE, fill = D_YEAR)) +
  geom_violin(alpha = 0.5, trim = FALSE, bw = 0.5) +
  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) + 
  geom_point(data = df_grade_means, aes(y = GAD_mean), shape = 18, size = 4, color = "black") +
  geom_line(data = df_grade_means, aes(y = GAD_mean, group = 1), color = "black", linewidth = 1) +
  theme_minimal() +
  labs(x = "Grade Level", y = "GAD-7 Score",
       title = "GAD-7 Scores by Grade Level (All Years Combined)", subtitle = "Higher scores indicate greater anxiety severity. Black Line = Mean Score per Grade") +
  theme(legend.position = "none",
        plot.subtitle = element_text(size = 10))


print(p_phq_all)
print(p_burn_all)
print(p_gad_all)


ggsave("PHQ_GradeLevel_AllYears.png", plot = p_phq_all, width = 6, height = 6, units = "in", dpi = 300, bg = "white")
ggsave("Burnout_GradeLevel_AllYears.png", plot = p_burn_all, width = 6, height = 6, units = "in", dpi = 300, bg = "white")
ggsave("GAD_GradeLevel_AllYears.png", plot = p_gad_all, width = 6, height = 6, units = "in", dpi = 300, bg = "white")

```


```{r most recent plotting}



df_recent <- df_allYears %>% 
  filter(Year == "2025-2026") %>%
  mutate(D_YEAR = factor(D_YEAR)) 


df_recent_means <- df_recent %>%
  group_by(D_YEAR) %>%
  summarise(
    PHQ_mean = mean(PHQ_SCORE, na.rm = TRUE),
    BUR_mean = mean(BUR_SCORE, na.rm = TRUE),
    GAD_mean = mean(GAD_SCORE, na.rm = TRUE)
  )


p_phq_recent <- ggplot(df_recent, aes(x = D_YEAR, y = PHQ_SCORE, fill = D_YEAR)) +
  geom_violin(alpha = 0.5, trim = FALSE, bw = 0.5) +
  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) + 
  geom_point(data = df_recent_means, aes(y = PHQ_mean), shape = 18, size = 4, color = "black") +
  geom_line(data = df_recent_means, aes(y = PHQ_mean, group = 1), color = "black", linewidth = 1) +
  theme_minimal() +
  labs(x = "Grade Level", y = "PHQ-9 Score",
       title = "2025-2026 PHQ-9 by Grade Level", subtitle = "Higher scores indicate greater depression severity. Black Line = Mean Score per Class") +
  theme(legend.position = "none",
        plot.subtitle = element_text(size = 10))


p_burn_recent <- ggplot(df_recent, aes(x = D_YEAR, y = BUR_SCORE, fill = D_YEAR)) +
  geom_violin(alpha = 0.5, trim = FALSE, bw = 0.5) +
  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) + 
  geom_point(data = df_recent_means, aes(y = BUR_mean), shape = 18, size = 4, color = "black") +
  geom_line(data = df_recent_means, aes(y = BUR_mean, group = 1), color = "black", linewidth = 1) +
  theme_minimal() +
  labs(x = "Grade Level", y = "Burnout Score",
       title = "2025-2026 Burnout by Grade Level", subtitle = "Higher scores indicate greater burnout severity. Black Line = Mean Score per Class") +
  theme(legend.position = "none",
        plot.subtitle = element_text(size = 10))


p_gad_recent <- ggplot(df_recent, aes(x = D_YEAR, y = GAD_SCORE, fill = D_YEAR)) +
  geom_violin(alpha = 0.5, trim = FALSE, bw = 0.5) +
  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) + 
  geom_point(data = df_recent_means, aes(y = GAD_mean), shape = 18, size = 4, color = "black") +
  geom_line(data = df_recent_means, aes(y = GAD_mean, group = 1), color = "black", linewidth = 1) +
  theme_minimal() +
  labs(x = "Grade Level", y = "GAD-7 Score",
       title = "2025-2026 GAD-7 by Grade Level", subtitle = "Higher scores indicate greater anxiety severity. Black Line = Mean Score per Class") +
  theme(legend.position = "none",
        plot.subtitle = element_text(size = 10))


print(p_phq_recent)
print(p_burn_recent)
print(p_gad_recent)


ggsave("PHQ_Recent_Square.png", plot = p_phq_recent, width = 6, height = 6, units = "in", dpi = 300, bg = "white")
ggsave("Burnout_Recent_Square.png", plot = p_burn_recent, width = 6, height = 6, units = "in", dpi = 300, bg = "white")
ggsave("GAD_Recent_Square.png", plot = p_gad_recent, width = 6, height = 6, units = "in", dpi = 300, bg = "white")

```


## SUMMARY (IMPORTANT)

We specifically look at two cohorts, the class of 2026 and 2027. Both the classes had significant negative squared terms for study year for the PHQ and GAD scores*, and significant positive terms for the non squared terms study year. The class of 2026 did not have a statistically significant coefficients for the burnout score, but class of 2027 did, and they also had a negative squared term and positive non squared term.

* note, one significant result (class of 2026 phq squared term) was not significant after doing anti-p-hacking adjustments

TLDR: - people got much worse last year - this year the two cohorts  were a bit better than last year in terms of depression (phq) and anxiety (gad), but the scores are still worse than the first year.
Evidence for the same pattern for burnout was only found for the class of 2027.


Here are the results from the models

ANALYSIS FOR: Class of 2026 

PHQ Coefficients:
              Estimate Std. Error t value Pr(>|t|)    
(Intercept)     4.3103     0.7132   6.044 6.56e-08 ***
Year_Num        4.8261     1.8154   2.658  0.00972 **  SIGNIF
I(Year_Num^2)  -2.0115     0.9463  -2.126  0.03706 *   
(note, after correcting for the familiy wise error rate)

BUR Coefficients:
              Estimate Std. Error t value Pr(>|t|)    
(Intercept)     1.4483     0.1490   9.722 1.27e-14 ***
Year_Num        0.4526     0.3792   1.194    0.237    
I(Year_Num^2)  -0.2134     0.1976  -1.079    0.284    

GAD Coefficients:
              Estimate Std. Error t value Pr(>|t|)    
(Intercept)     4.1034     0.6449   6.363 1.77e-08 ***
Year_Num        4.7615     1.6416   2.900  0.00498 ** SIGNIF
I(Year_Num^2)  -2.1149     0.8557  -2.472  0.01589 *  SIGNIF





ANALYSIS FOR: Class of 2027 

PHQ Coefficients:
              Estimate Std. Error t value Pr(>|t|)    
(Intercept)     5.4000     0.8206   6.580 8.97e-09 ***
Year_Num        9.6516     2.1549   4.479 3.05e-05 *** SIGNIF
I(Year_Num^2)  -4.7472     1.0492  -4.525 2.59e-05 *** SIGNIF

BUR Coefficients:
              Estimate Std. Error t value Pr(>|t|)    
(Intercept)     1.2800     0.1549   8.264 8.92e-12 ***
Year_Num        1.6659     0.4067   4.096 0.000117 *** SIGNIF
I(Year_Num^2)  -0.7720     0.1980  -3.899 0.000229 *** SIGNIF

GAD Coefficients:
              Estimate Std. Error t value Pr(>|t|)    
(Intercept)     4.8000     0.9205   5.215 1.99e-06 ***
Year_Num       13.3124     2.4170   5.508 6.43e-07 *** SIGNIF
I(Year_Num^2)  -6.6776     1.1768  -5.674 3.36e-07 *** SIGNIF

```{r first vs last}

cat("\n=================================================\n")
cat("FIRST vs LAST YEAR COMPARISON \n")
cat("=================================================\n")

df_first_last <- df_model %>%
  filter(Year_Num %in% c(0, 2))

cohorts <- unique(df_first_last$Cohort)

for (c in c("Class of 2026","Class of 2027")) {
  cat("\n----------------------------------\n")
  cat("ANALYSIS FOR:", c, "\n")
  sub_data <- df_first_last %>% filter(Cohort == c)
  
  if (nrow(sub_data) > 0 && length(unique(sub_data$Year_Num)) > 1) {
    print("--PHQ (First vs Last)--")
    print(summary(lm(PHQ_SCORE ~ Year_Num, data = sub_data)))
    
    print("--BUR (First vs Last)--")
    print(summary(lm(BUR_SCORE ~ Year_Num, data = sub_data)))
    
    print("--GAD (First vs Last)--")
    print(summary(lm(GAD_SCORE ~ Year_Num, data = sub_data)))
  } else {
    cat("Insufficient data for this cohort (needs both Year 0 and Year 2).\n")
  }
}

for (c in c("Class of 2026","Class of 2027")) {
  cat("\n----------------------------------\n")
  cat("ANALYSIS FOR:", c, "\n")
  sub_data <- df_first_last %>% filter(Cohort == c)
  
  if (nrow(sub_data) > 0 && length(unique(sub_data$Year_Num)) > 1) {
    print("--PHQ (First vs Last)--")
    print(summary(lm(PHQ_SCORE ~ Year_Num + D_YEAR, data = sub_data)))
    
    print("--BUR (First vs Last)--")
    print(summary(lm(BUR_SCORE ~ Year_Num + D_YEAR, data = sub_data)))
    
    print("--GAD (First vs Last)--")
    print(summary(lm(GAD_SCORE ~ Year_Num + D_YEAR, data = sub_data)))
  } else {
    cat("Insufficient data for this cohort (needs both Year 0 and Year 2).\n")
  }
}
```